---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
options(max.print = 60)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# neuromat

<!-- badges: start -->
[![R-CMD-check](https://github.com/capro-uio/neuromat/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/capro-uio/neuromat/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/neuromat)](https://CRAN.R-project.org/package=neuromat)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/capro-uio/neuromat/branch/main/graph/badge.svg)](https://app.codecov.io/gh/capro-uio/neuromat?branch=main)
<!-- badges: end -->

Running vertex-wise linear mixed effects models in [Freesurfer](https://surfer.nmr.mgh.harvard.edu/fswiki/LinearMixedEffectsModels) requires a specific file format. 
The qdec format requires neatly created binary columns for 
categorical value levels, and is made more efficient if continuous
variables are scaled (or de-meaned). 
This package provides functions to create these files from R.


## Installation

You can install the development version of `neuromat` like so:

``` r
# Install Capro R-universe
install.packages('neuromat', 
  repos = 'https://capro-uio.r-universe.dev')
```

## Example

```{r}
library(neuromat)
cars <- mtcars
cars$cyl <- factor(cars$cyl)
cars$gear <- factor(cars$gear)
qdec <- make_fs_qdec(cars, mpg ~ -1 + cyl + hp)
plot(qdec)
```
